---
title: "GallupDataPreparation"
author: Phyllis Lun
date: '2025-06-18'
---

## 0. Subset the Gallup raw data

The version of the Gallup raw data used in this study is "The_Gallup_022125.dat". 

This data was released to the [UofT Map and Data Library](https://mdl.library.utoronto.ca/collections/numeric-data/gallup) in February 2024.

Since the raw data file is relatively large, we used a separate environment to run the following code, which subsets the variables that we needed.

```{r eval=FALSE}
# The code in this chunk will not be run in the current Rmd
# load "dplyr"
library(dplyr)
library(data.table)

```{r eval=FALSE}
# This code will not be run in the current Rmd

# Import the data file "The_Gallup_022125.dat"
gallupRaw <- rio::import("~/Documents/GitHub/DATA/The_Gallup_022125.dat")

# Create a character vector of the names of the variables we're interested in
varRequired <- c(
  "COUNTRYNEW","EMP_2010", "HHSIZE", "INCOME_1", "INCOME_2", "INCOME_3", "INCOME_4", "INCOME_5",
  "MODE", "REGION_MMR", "WGT", "WP10251", "WP103", "WP108", "WP109", "WP110", "WP111", "WP112",
  "WP119", "WP1219", "WP1220", "WP1223", "WP1233", "WP13125",
  "WP1325", "WP134", "WP137", "WP138", "WP139", "WP14", "WP144", "WP145", "WP146",
  "WP150", "WP16", "WP18", "WP3117", "WP4", "WP40", "WP43", "WP63", "WP67", "WP69",
  "WP70", "WP71", "WP74", "WP98", "WPID", "YEAR_CALENDAR", "YEAR_WAVE"
)

# Subset gallupRaw to include only the variables specified in varRequired
gallupSubset <- gallupRaw %>% select(all_of(varRequired))

# Save the gallupSubset data frame to a .rds file for later use
saveRDS(gallupSubset, "(250703)GWP_RawSubset_MyanmarProject.rds")
```


## 1. Clean the Subset Gallup Data

### 1.1 Import the subset raw data

Once we had the subset of the Gallup data, we imported the data and loaded the necessary packages:

```{r set-up}
# Load "dplyr"
library(dplyr)

# Read the RDS file
gallupSubset <- readRDS("data/(250618)GWP_RawSubset_MyanmarProject.rds")
```


### 1.2 Create `YEAR_INTERVIEW` 

The existing 'YEAR_CALENDAR' and 'YEAR_WAVE' variables may not accurately represent the year a participant was interviewed. 'YEAR_CALENDAR' is based on the last interview day for all participants, while 'YEAR_WAVE' refers to the wave year. We extracted the interview year from 'WP4:Interview Date' and created a 'YEAR_INTERVIEW' variable. If the 'Interview Date' was missing, we used 'YEAR_CALENDAR' as 'YEAR_INTERVIEW'.

```{r creating-YEAR_INTERVIEW}
gallup <- gallupSubset %>%
  # Convert WP4 to a date
  mutate(WP4 = as.Date(WP4, "%m/%d/%Y")) %>%
  # Create YEAR_INTERVIEW variable
  mutate(YEAR_INTERVIEW =
           ifelse(is.na(WP4), YEAR_CALENDAR, format(WP4, format = "%Y")))
```

### 1.3 Clean 'Don't Know' and 'Refused' Responses

Gallup coded some responses as 'Don't Know' or 'Refused', which would not be meaningful for our analysis. We converted these responses to NA using the values found in the [WORLD POLL REFERENCE TOOL](https://wpr.gallup.com/).

```{r converting-responses-to-NA}

# Convert certain values to NA
gallup <- gallup %>%
  mutate(
         WP10251 = ifelse(WP10251 %in% c(3, 4), NA, WP10251),
         WP103 = ifelse(WP103 %in% c(3, 4), NA, WP103),
         WP108 = ifelse(WP108 %in% c(3, 4), NA, WP108),
         WP109 = ifelse(WP109 %in% c(3, 4), NA, WP109),
         WP110 = ifelse(WP110 %in% c(3, 4), NA, WP110),
         WP111 = ifelse(WP111 %in% c(3, 4), NA, WP111),
         WP112 = ifelse(WP112 %in% c(3, 4), NA, WP112),
         WP119 = ifelse(WP119 %in% c(3, 4), NA, WP119),
         WP1220 = ifelse(WP1220 %in% c(100), NA, WP1220),
         WP1223 = ifelse(WP1223 %in% c(6, 7), NA, WP1223),
         WP1233 = ifelse(WP1233 %in% c(98, 99, 97), NA, WP1233),
         WP13125 = ifelse(WP13125 %in% c(3, 4), NA, WP13125),
         WP1325 = ifelse(WP1325 %in% c(3, 4), NA, WP1325),
         WP134 = ifelse(WP134 %in% c(3, 4), NA, WP134),
         WP137 = ifelse(WP137 %in% c(3, 4), NA, WP137),
         WP138 = ifelse(WP138 %in% c(3, 4), NA, WP138),
         WP139 = ifelse(WP139 %in% c(3, 4), NA, WP139),
         WP14 = ifelse(WP14 %in% c(4, 5), NA, WP14),
         WP144 = ifelse(WP144 %in% c(3, 4), NA, WP144),
         WP145 = ifelse(WP145 %in% c(3, 4), NA, WP145),
         WP146 = ifelse(WP146 %in% c(3, 4), NA, WP146),
         WP150 = ifelse(WP150 %in% c(3, 4), NA, WP150),
         WP16 = ifelse(WP16 %in% c(98, 99), NA, WP16),
         WP18 = ifelse(WP18 %in% c(98, 99), NA, WP18),
         WP3117 = ifelse(WP3117 %in% c(4, 5), NA, WP3117),

         WP40 = ifelse(WP40 %in% c(3, 4), NA, WP40),
         WP43 = ifelse(WP43 %in% c(3, 4), NA, WP43),
         WP63 = ifelse(WP63 %in% c(3, 4), NA, WP63),
         WP67 = ifelse(WP67 %in% c(3, 4), NA, WP67),
         WP69 = ifelse(WP69 %in% c(98, 99), NA, WP69),
         WP70 = ifelse(WP70 %in% c(98, 99), NA, WP70),
         WP71 = ifelse(WP71 %in% c(98, 99), NA, WP71),
         WP74 = ifelse(WP74 %in% c(3, 4), NA, WP74),
         WP98 = ifelse(WP98 %in% c(3, 4), NA, WP98))


### 1.4 Convert Binary Response Variables

For questions that have binary responses, such as 'WP119' which asks "Is religion an important part of your daily life?", the responses were initially coded as 1 for 'Yes' and 2 for 'No'. However, for regression analysis, it is more common and interpretable to code binary variables as 0 and 1. Therefore, we converted the responses for binary questions from 1 and 2 to 0 and 1, respectively.

```{r converting-binaryQ}
# Convert "2" to "0"
gallup <- gallup %>%
  mutate(
    WP10251 = ifelse(WP10251 == 2, 0, WP10251),
    WP103 = ifelse(WP103 == 2, 0, WP103),
    WP108 = ifelse(WP108 == 2, 0, WP108),
    WP109 = ifelse(WP109 == 2, 0, WP109),
    WP110 = ifelse(WP110 == 2, 0, WP110),
    WP111 = ifelse(WP111 == 2, 0, WP111),
    WP112 = ifelse(WP112 == 2, 0, WP112),
    WP119 = ifelse(WP119 == 2, 0, WP119),
    WP137 = ifelse(WP137 == 2, 0, WP137),
    WP138 = ifelse(WP138 == 2, 0, WP138),
    WP139 = ifelse(WP139 == 2, 0, WP139),
    WP144 = ifelse(WP144 == 2, 0, WP144),
    WP145 = ifelse(WP145 == 2, 0, WP145),
    WP146 = ifelse(WP146 == 2, 0, WP146),
    WP150 = ifelse(WP150 == 2, 0, WP150),
    WP40 = ifelse(WP40 == 2, 0, WP40),
    WP43 = ifelse(WP43 == 2, 0, WP43),
    WP63 = ifelse(WP63 == 2, 0, WP63),
    WP67 = ifelse(WP67 == 2, 0, WP67),
    WP69 = ifelse(WP69 == 2, 0, WP69),
    WP70 = ifelse(WP70 == 2, 0, WP70),
    WP71 = ifelse(WP71 == 2, 0, WP71),
    WP74 = ifelse(WP74 == 2, 0, WP74),

    WP103 = ifelse(WP103 == 2, 0, WP103),
    WP13125 = ifelse(WP13125 == 2, 0, WP13125),
    WP1325 = ifelse(WP1325 == 2, 0, WP1325),
    WP98 = ifelse(WP98 == 2, 0, WP98),
    WP1219 = ifelse(WP1219 == 2, 0, WP1219),
    WP134 = ifelse(WP134 == 2, 0, WP134)
  )
```


The object `gallup` is now a cleaned subset of the Gallup data.

We save the Gallup cleaned data to a .rds file for later use

```{r exporting-gallup-cleaned}
saveRDS(gallup, "data/(250701)GWP_cleaned_MyanmarProject.rds")
```


